---
title: "Lab 2- Spatial Statistics"
author: "Harish Saifi"
date: "2024-08-24"
output:
  pdf_document: default
  html_document: default
---
Objective : 
(i) Generate a set of 10 points without attributes on the unit square [0, 1] × [0, 1] from uniform distribution and convert them into spatial points, make a scattered plot of them in R .

(ii)  Also, create a Spatial Points Data frame object by building it from a Spatial Points object and a data.frame containing the attributes, and plot these spatial points in coordinate system.

In this lab we are going to transform our simple coordinates data into spatial data by usig the `sp` library.

```{r}
#install.packages("sp")
library(sp)
```
# 1. Generating Random Coordinates:

xc and yc create vectors of 10 random numbers each, rounded to two decimal places. These vectors represent x and y coordinates. xy combines these vectors into a matrix of coordinate pairs
```{r}
xc=round(runif(10),2)
yc=round(runif(10),2)
xy=cbind(xc,yc)
```
# 2. Creating and Plotting Spatial Points:

xy.sp converts the xy matrix into a SpatialPoints object, which is a spatial data type in R. plot(xy.sp, pch=7) plots these spatial points using a specific symbol (pch=7).
```{r}
xy.sp=SpatialPoints(xy)
xy.sp[1:3,]
plot(xy.sp,pch=7)
```
# 3. Creating and Plotting Spatial Points:

xy.sp converts the xy matrix into a SpatialPoints object, which is a spatial data type in R. plot(xy.sp, pch=7) plots these spatial points using a specific symbol (pch=7).
```{r}
xy.cc=coordinates(xy.sp)
class(xy.cc)
```
# 4. Extracting Coordinates and Checking Class:

xy.cc extracts the coordinates from the SpatialPoints object. class(xy.cc) checks the class of the extracted coordinates.
```{r}
bbox(xy.sp)
dimensions(xy.sp)
xy.sp[1:2]
xy.df=as.data.frame(xy.sp)
class(xy.df)
dim(xy.df)
summary(xy.sp)
```
# 5. Getting Bounding Box, Dimensions, and Summary:

bbox(xy.sp) retrieves the bounding box of the spatial points, defining the area covering all points.
## 2 Points with attributes

```{r}
df=data.frame(z1=round(5+rnorm(10),2),z2=20:29)
df[1:3,]
xy.sp.df=SpatialPointsDataFrame(xy.sp,df)
xy.sp.df[1:3,]
summary(xy.sp.df)
dimensions(xy.sp.df)
xy.sp.df[1:2,]
xy.sp.df[1][1:3,]
xy.sp.df[1:2,"z2"]


xy.df=as.data.frame(xy.sp.df)
xy.df[1:3,]
xy.cc=coordinates(xy.sp.df)
class(xy.cc)
dim(xy.cc)
df1=data.frame(xy,df)
coordinates(df1)=c("xc","yc")
df1[1:3,]
df2=data.frame(xy,df)
coordinates(df2)=~xc+yc
df2[1:3,]
as.data.frame(df2)[1:2,]
coordinates(df2)[1:2,]
df2[["Z2"]]
df2[["Z2"]][10]=20
df2[["z2"]]=1:10
summary(df2)
```
# 6. Visualizing Data with Bubble and Spplot:

bubble(df2, "z1", key.space="bottom") creates a bubble plot of the df2 data, using z1 as the variable for bubble sizes. spplot(df2, "z1", key.space="bottom") creates a spatial plot of df2, also using z1 for visualization, with a legend placed at the bottom.
```{r}
bubble(df2,"z1",key.space="left")
spplot(df2,"z1",key.space="right")

```


