---
title: "Lab 1- Spatial Statistics"
author: "HARISH SAIFI (GL6134/23DSMSA112)"
date: "2024-08-30"
output: 
  pdf_document: default
  html_document: default
---
\vspace{-0.5cm}
# 1. Installing and Loading the imager Package:
```{r message=FALSE}
#install.packages("imager")
library(imager)
```
# 2.Plotting the boats Image:

The plot(boats) function displays the boats image. The boats variable is a sample image provided by the imager package. Simply typing boats will show information about the image object
```{r}
plot(boats);boats
```

# 3. Converting the Image to Grayscale:

The grayscale(boats) function converts the boats image to grayscale. This process removes the color information, leaving only shades of gray, which simplifies the image
```{r}
grayscale(boats)
```
# 4. Performing Image Transformations:

Mathematical operations are applied to the boats image:
z1 involves taking the logarithm of the image and adding three times the square root of the image.
z2 adds three times the square root of the image to itself.
These operations create transformed versions of the image, which are then displayed.

```{r}
z1=log(boats)+3*sqrt(boats)
z2=boats+3*sqrt(boats)
z1;z2
```

The layout(t(1:3)) function configures the plotting area to show three plots in a single row.
```{r}
par(mfrow=c(1,2));plot(z1);plot(z2)

layout(t(1:3));plot(boats);plot(boats/2);plot(boats,reshape=FALSE)
```

# 6. Creating a Histogram of Grayscale Values:

The grayscale version of the boats image is used to create a histogram that displays the distribution of luminescence (brightness) values in the image. The main parameter adds a title to the histogram.

```{r}
grayscale(boats)%>%hist(main="Luminance value in boats pictures")
```
\vspace{-0.5cm}
